---
title: "Animation ndvi Portugal 2017"
author: "William Martinez"
date: "November 20, 2018"
output: 
  html_document:
    theme: journal
    toc: true
    toc_depth: 4
    toc_float: true
runtime: shiny
---

<style>
body {
text-align: justify}
</style>


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(kableExtra)
```

In the frame of a multitemporal analysis of composites of sentinel 2 imagery, I want to document how important is to make dynamic the training data in the classification. According with the explorative analysis, we visualize trajectories of ndvi, so that we  perceived how some land cover with vegetation tend to show different ndvi ranges depending on the time of the year. However, we also saw how some trajectories of ndvi for some classes, such as water or no- vegetation, that suppose to be static over the year, depict strong variance and ranges of ndvi that teoritically are not supported. Well, I will go through different examples showing why we are having these unusual patterns by using 4 compositions of NDVI for 4 different seasons of the year. I expect ndvi can be determinant in the rename of some labels, and therefore, we can define a new workflow for the multitemporal classification.


### Estuaries 

My first example here is Estuaries. I remember that the water labels were subcategorized into eight classes and then filtered into three classes that I think are the best representation of class water. Therefore, in the next figure we can visualize how ndvi differs in space and time in the conection of Tagus river with the atlantic Ocean. In this sense, I made a query only conssidering sample points of Estuaries, so that, we can visualize how some points do not overlap with water during certain times of the year. Since the near infrared reflects less than the red for water, I defined points intersected by pixels with ndvi lower than zero as water samples and larger as temporal samples of non - vegetation.

```{r estuaries, echo=FALSE, message=FALSE}
library("shinyWidgets")
library(tibble)
library(sf)
library(rgdal)
library(raster)
library(ggplot2)
#seasons of the imagery
se = c("Spring","Summer","Autumn", "Winter")
seasons = factor(se, levels = se)

inputPanel(
      sliderTextInput(
               inputId = "a",
               label = "Season range slider:",
               choices = seasons,
               selected = seasons[1])
)

renderPlot({
  #selection
  ind = which(input$a==seasons)
  #shapefile
  shape_file = "./shp/water_class2.shp"
  points_water = sf::read_sf(shape_file)
  #generate plots based on input$a 
  dir_raster = "./data"
  file_raster = list.files(dir_raster)
  r = vector("list",4)
  j = 1
  for(i in file_raster){
    if(endsWith(i, "clip1.tif"))
    {
      r[[j]] = raster(paste0(dir_raster,"/",i))
      j = j + 1
    }
  }
  selec_col = points_water[,input$a]
  st_geometry(selec_col) = NULL
  colnames(selec_col) = c("axis")
  color_points = bg = c("red", "blue")[unclass(as.factor(selec_col$axis))]
  plot(r[[ind]],zlim = c(-1, 1), main = paste0("NDVI over Estuaries, date: ", input$a))
  plot(points_water[,1], add= TRUE, pch = 16, col= color_points , cex = 0.5)
  legend("topright",legend=c("non-vegetated", "water") ,col=c("red","blue"),pch=16,bty="n",cex=1,pt.cex=1)
})
```


### Herbaceus periodic 


```{r herbaceus, echo=FALSE, message=FALSE}
inputPanel(
      sliderTextInput(
               inputId = "b",
               label = "Season range slider:",
               choices = seasons,
               selected = seasons[1])
)

renderPlot({
  #selection
  ind = which(input$b==seasons)
  #shapefile
  shape_file = "./shp/HerbaceusPeriodic.shp"
  points_hp = sf::read_sf(shape_file)
  #generate plots based on input$a 
  dir_raster = "./data"
  file_raster = list.files(dir_raster)
  r1 = vector("list",4)
  j = 1
  for(i in file_raster){
    if(endsWith(i, "clip2.tif"))
    {
      r1[[j]] = raster(paste0(dir_raster,"/",i))
      j = j + 1
    }
  }
  selec_col = points_hp[,"CLASS_NAME"]
  st_geometry(selec_col) = NULL
  colnames(selec_col) = c("axis")
  color_points = bg = c("red", "blue")[unclass(as.factor(selec_col$axis))]
  plot(r1[[ind]],zlim = c(0, 1), main = paste0("NDVI over Herbaceous periodic, date: ", input$b))
  plot(points_hp[,1], add= TRUE, pch = 16, col= color_points , cex = 1)
  legend("topright",legend=c("Irrigated", "Non_rirrigated") ,col=c("red","blue"),pch=16,bty="n",cex=1,pt.cex=1)
})

```

### Multitemporal training dataset

```{r stdataset}
file = "./shp/training_samples_cleaned.shp"
#file = "./shp/training_samples_composites_cleaned.shp"
data_ndvi = sf::read_sf(file)
#classes with vegetation
classes_names = c("Coniferous_trees", "Eucalyptus_trees","Holm_and_Cork_Trees","Bushes_and_shrubs", "Herbaceous_periodic",
                  "Natural_Herbaceous","Herbaceous_permanet","Rice_fields","Water","Wetlands","Sealed","Non_vegetated")
#query over previous class
query_veget = data_ndvi[data_ndvi$CLASS_NAME %in% classes_names[1:6] ,-1]
#geometry
geo = st_geometry(query_veget)
st_geometry(query_veget) =  NULL
#function to separate non-vegetation
vegetation = function(v, cl){
  y = v
  tr = 0.3
  y[v >= tr] = cl[v >= tr]
  y[v < tr] = "Non_vegetated"
  return(y)
}
list_veget = lapply(query_veget[,-1], vegetation, cl = query_veget$CLASS_NAME)
df_veget = data.frame(list_veget)
#Merging geometry to data frame
st_geometry(df_veget) = geo

#merging rest of the classes
df_rest = data_ndvi[!(data_ndvi$CLASS_NAME %in% classes_names[1:6]),-1]
geo2 = st_geometry(df_rest)
st_geometry(df_rest) = NULL
labeldf = function(x, cl) x[] = cl
list_rest = lapply(df_rest[,-1],labeldf, cl = df_rest$CLASS_NAME)
df_rest2 = data.frame(list_rest)
st_geometry(df_rest2) = geo2
#merging
df_final = rbind(df_veget,df_rest2)
#Names of the colums
names_images =  c('20170115', "20170405", "20170525", "20170604", "20170614",'20170704', '20170709', '20170714', '20170724', '20170729', '20170803', '20170808', '20170813', '20170818', '20170902', '20170907', '20170912', '20170922', '20170927', '20171002', '20171012', '20171022', '20171027', '20171111', '20171116', '20171121', '20171201', '20171216', '20171221',"geometry")
colnames(df_final) = names_images
#writing objet sf as shapefile
#file_out = './shp/temporal_dataset_composites.shp'
#sf::write_sf(df_final, dsn = file_out)
```




```{r stdataset2}
#removing geometry
st_geometry(df_final) = NULL
#Days of the images
day_shoot = c("2017/01/15", "2017/04/05", "2017/05/25", "2017/06/04", "2017/06/14",
              "2017/07/04", "2017/07/09", "2017/07/14", "2017/07/24", "2017/07/29",
              "2017/08/03", "2017/08/08", "2017/08/13", "2017/08/18", "2017/09/02" ,
              "2017/09/07", "2017/09/12", "2017/09/22", "2017/09/27", "2017/10/02",
              "2017/10/12", "2017/10/22", "2017/10/27", "2017/11/11", "2017/11/16",
              "2017/11/21", "2017/12/01","2017/12/16", "2017/12/21")
output = NULL
list_df_final = lapply(df_final,function(x) table(x))
j = 1
for(i in list_df_final){
  dft = data.frame(day_shoot[j],i)
  colnames(dft) = c("Date","Class","freq")
  output = rbind(output, dft)
  j = j + 1
}
#defining position of the text
output$halffreq = output$freq *0.5
output$pos = 1
for(i in output$Date){
  ind1 = which(output$Date == i)
  out1 = output[ind1,]
  out0 = 0
  for(j in classes_names){
    ind2 = which(out1$Class == j)
    out0 = out0 + out1[ind2,3]
    output[ind1[ind2],5] = dim(df_final)[1] - (out0 - output[ind1[ind2],4])
  }
}

output$Class = factor(output$Class, classes_names)
#colors in the graphic
colors_classes = c("green4", "green3", "green1", "springgreen3", "springgreen1", "darkolivegreen3", "turquoise","plum2","blue2","royalblue2", "gray", "goldenrod")
names(colors_classes) = classes_names

ggplot(data=output, aes(x=Date, y=freq, fill= Class)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
  geom_bar(stat="identity") +
  scale_fill_manual("Classes", values = colors_classes) +
  geom_text(aes(label = freq,y = pos),size = 2)

```










